---
title: "Project4"
author: "Jan Troeltsch"
date: "25.01.2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(tidylog)
library(ggplot2)
library(lubridate)
library(stringr)
library(readxl)
library(datapasta)
library(DBI)
library(RSQLite)
library(devtools)
library(encryptr)
library(ggmap)
library(magrittr)
library(visdat)
library(infer)
library(readr)
library(knitr)
library(kableExtra)
```

```{r message=FALSE, warning=FALSE, include=FALSE}

```
# Exercises on Swiss Cows Data
## Part 1: Import the data

With the following lines of Code we import the data from a csv file
```{r message=FALSE, warning=FALSE}
cows <- read_csv("data/swiss_cows_data.csv")
```
Here is the complete data:
    
```{r echo=FALSE}
cows %>% kable() %>% 
kable_minimal(full_width = F, position = "center")
```

## Part 2: Perform a t-test using the first 3 cows of each breed

```{r message=FALSE, warning=FALSE, include=FALSE}
cows_03 <- cows %>% filter(num <= 3)
```
Before doing a t-test on the data for the first 3 cows, I produced the following summary table with sample size, arithmetic mean, and standard deviation for each breed of cattle:
    
```{r echo=FALSE, message=FALSE, warning=FALSE}
cows_03_summary <- cows_03 %>% 
  group_by(breed) %>% 
  transmute(breed,mean(weight),median(weight),sd(weight)) %>% 
  distinct()
cows_03_summary%>% 
  kable() %>% 
  kable_minimal(full_width = F, position = "center")
```
Based on these results I was able to answer the first question.

Then I used the t_test function from the {infer} package to get an objective comparison between the two obtained means. The result looks as follows:
```{r echo=FALSE, message=FALSE, warning=FALSE}
my_test <- cows_03 %>%
  t_test(formula = weight ~ breed, order = c("Brown_Swiss","Swiss_Red_Holstein"),conf_level = 0.95)
my_test %>% 
  kable() %>% 
  kable_minimal(full_width = F, position = "center")
```

As the interval between the lower_ci and the upper_ci contains the zero value, the cows_03 data set does not provide enough evidence that Brown Swiss are heavier than Swiss Red Holstein within our desired 95% confidence level. This is also confirmed by the p-value of 0.279: having chosen a confidence level of 0.95, and the resulting p-value being higher than 0.05, we conclude that the difference between the two average weights is not statistically significant at the level of 0.05.

With the following code we can visualize the steps of the t-test process
```{r echo=FALSE, message=FALSE, warning=FALSE}
t_obs <- my_test %>%
  select(statistic)

cows_03 %>%
  specify(formula = weight ~ breed) %>%
  hypothesize(null = 'independence') %>%
  calculate(stat = "t") %>%
  visualise(method = "theoretical") +
  shade_p_value(obs_stat = t_obs, direction = "both") +
  theme_minimal()
```

The area of the red region under the density curve corresponds to the p-value of 0.279

## Part 3: Perform a t-test using only the first cow of each breed

Performing a t-test on just the data of the first cow of each breed does give an error, as there is not enough data. Here is the error message:
```{r echo=FALSE, error=TRUE, message=FALSE, warning=TRUE}
cows_01 <- cows %>% filter(num <= 1)

my_test <- cows_01 %>%
  t_test(formula = weight ~ breed, order = c("Brown_Swiss","Swiss_Red_Holstein"),conf_level = 0.95)
```

## Part 4: Perform a t-test using all the cows of each breed

Here is the result of the t-test on the data of all the cows:
```{r echo=FALSE}
my_test <- cows %>%
  t_test(formula = weight ~ breed, order = c("Brown_Swiss","Swiss_Red_Holstein"),conf_level = 0.95) 
my_test %>%
  kable() %>% 
  kable_minimal(full_width = F, position = "center")
```

And here is the density graph:

```{r echo=FALSE, message=FALSE, warning=FALSE}
cows %>%
  specify(formula = weight ~ breed) %>%
  hypothesize(null = 'independence') %>%
  calculate(stat = "t") %>%
  visualise(method = "theoretical") +
  shade_p_value(obs_stat = t_obs, direction = "both") +
  theme_minimal()
```

### Conclusion
The confidence interval [lower_ci = 42 ; upper_ci = 69] does not contain the zero value. The cows data set provides therefore enough evidence that Brown Swiss are heavier than Swiss Red Holstein.
The p-value is 0 and therefore it is 100% certain the differences in weight depending on breed are  statistically significant.

------
# Exercises on Fitness Club data

## Part 5: Import the data
Import in R the data from files fitness_members.csv and fitness_tracking.csv.
With the following lines of Code we import the data from a csv file
```{r message=FALSE, warning=FALSE}
fitness_members <- read_csv("data/fitness_members.csv")
fitness_tracking <- read_csv("data/fitness_tracking.csv")
```
Here are the first 10 rows of the fitness_members data:

```{r echo=FALSE}
fitness_members %>% 
  filter(id <="000010") %>% 
  kable() %>% 
kable_minimal(full_width = F, position = "center")

```
Here are the first 10 rows of the fitness_tracking data:
```{r echo=FALSE}
fitness_tracking %>% 
  filter(id <="000010") %>% 
  kable() %>% 
kable_minimal(full_width = F, position = "center")

```

## Part 6: t-test

In order to perform a T-Test to answer the question whether we observe a different evolution of BMI from week 0 to week 12 in the different fitness club membership categories, I need to first transform the data.
```{r include=FALSE}
fitness_members <- fitness_members %>% 
  mutate(BMI=weight/((height/100)*(height/100)))

fitness_tracking_long <- fitness_tracking %>% 
  full_join(fitness_members, by = c("id")) %>% 
  rename(wk_000=weight) %>% 
  select(id,starts_with("wk_")) %>% 
  
  pivot_longer(cols = c(starts_with("wk_")),
               names_to = "week",
               values_to = "weight") %>% 
  filter(!is.na(weight))

fitness_tracking_long <- fitness_tracking_long %>%
  left_join(fitness_members,by=c("id")) %>% 
  select(id,week,weight.x,height) %>% 
  rename(weight=weight.x) %>% 
  mutate(BMI=round(weight/((height/100)*(height/100)),digits=1),wk=as.numeric(substr(week,4,6))) %>% 
  select(id,wk,weight,BMI) %>% 
  arrange(id,wk) %>% 
  mutate(BMI_percent_change_vs_prev_week=((BMI/lag(BMI))-1)*100)
  
fitness_tracking_long_wk12 <- fitness_tracking_long %>% 
  filter(wk==12) %>% 
  transmute(id,BMI_wk12 = BMI)

fitness_members_bmi_evo_wk0_to_wk12 <- fitness_members %>% 
  left_join(fitness_tracking_long_wk12, by=c("id")) %>% 
  mutate(BMI_evo = BMI_wk12/BMI-1) %>% 
  filter(!is.na(BMI_wk12))
```
### Economy vs Balance
The t-test for comparing the catergories Economy and Balance does show a statistically significant difference in BMI reduction, but as the lower ci and uper ci are insignificantly low, this statistical difference is practically not relevant.
```{r echo=FALSE, message=FALSE, warning=FALSE}
my_test_eco_vs_bal <- fitness_members_bmi_evo_wk0_to_wk12 %>%
  filter(m_category!="Premium") %>% 
  t_test(formula = BMI_evo ~ m_category, order = c("Economy","Balance"),conf_level = 0.99) 
my_test_eco_vs_bal %>%
  kable() %>% 
  kable_minimal(full_width = F, position = "center")
```

### Balance vs Premium
The t-test for comparing the catergories Balance and Premium does show a statistically significant difference in BMI reduction, but as the lower ci and uper ci are insignificantly low, this statistical difference is practically not relevant.
```{r echo=FALSE, message=FALSE, warning=FALSE}
my_test_bal_vs_prem <- fitness_members_bmi_evo_wk0_to_wk12 %>%
  filter(m_category!="Economy") %>% 
  t_test(formula = BMI_evo ~ m_category, order = c("Balance","Premium"),conf_level = 0.99) 
my_test_bal_vs_prem %>%
  kable() %>% 
  kable_minimal(full_width = F, position = "center")
```

### Economy vs Premium
The t-test for comparing the catergories Economy and Premium does show a statistically significant difference in BMI reduction, but as the lower ci and uper ci are insignificantly low, this statistical difference is practically not relevant.
```{r echo=FALSE, message=FALSE, warning=FALSE}
my_test_eco_vs_prem <- fitness_members_bmi_evo_wk0_to_wk12 %>%
  filter(m_category!="Balance") %>% 
  t_test(formula = BMI_evo ~ m_category, order = c("Economy","Premium"),conf_level = 0.99) 
my_test_eco_vs_prem %>%
  kable() %>% 
  kable_minimal(full_width = F, position = "center")
```

## Part 7: Chi-squared test


```{r echo=FALSE, message=FALSE, warning=FALSE}
fitness_members <-  fitness_members %>%
  mutate(m_category = factor(m_category, levels = c("Economy", "Balance", "Premium")))
```

I used the chisq_test() function of the {infer} package to test the independence of member category and gender. Here is the result:
```{r echo=FALSE, message=FALSE, warning=FALSE}
chi_test <- chisq_test(fitness_members,formula = m_category ~ gender)
chi_test %>% 
  kable() %>% 
  kable_minimal(full_width = F, position = "center")
```
As the p-value is 'large' we can conclude that member category and gender are rather independent.


## Part 8:  Visualise p-value

I visualised the p-value with the following code:
```{r echo=TRUE, message=FALSE, warning=FALSE}
fitness_members %>%
  specify(formula = m_category ~ gender) %>%
  hypothesize(null = 'independence') %>%
  calculate(stat = "Chisq") %>%
  visualise(method = "theoretical") +
  shade_p_value(obs_stat = t_obs, direction = "right") +
  theme_minimal()
```

The red shaded area shows the p-value.


***
End of Report
